---
title: "Covariate Creation"
output:
  html_document: default
  html_notebook:
    number_sections: yes
    toc: yes
  pdf_document: default
---
#Covariates

Covariates are the same as features, x variables. 


# Head
```{r, message=FALSE, warning=FALSE}
#clear
rm(list=ls())
#packages
library(caret)
library(ISLR)
library(dplyr)
library(splines)

```
# Data
```{r}
data(Wage) 
colnames(Wage)
```
## Training and testing
```{r}
inTrain <- createDataPartition(y=Wage$wage,list = FALSE,p = .75)
training <- Wage[inTrain,]
testing <- Wage[-inTrain,]
```

# Covariates tidying and adding
Tidy covariates = making covariates from existing covariates. Needed more for regression and svms, less so for classification trees. 

Should only be done on the training set. 

## Create dummies from categoricals
```{r}
table(training$jobclass)
dummies <- dummyVars(wage ~ jobclass,data=training,fullRank = TRUE)
predict(dummies,newdata=training) %>% head
```

## Remove covariates with little variation

```{r}
nearZeroVar(training,saveMetrics = TRUE)
training2 <- training[,!nsv$nzv]
nearZeroVar(training2,saveMetrics = TRUE)

```

## Spline basis

For fitting curvey lines to data. BS function creates a 3rd degree polynomial. Columns represent x, x^2, x^3. 

```{r}
bsBasis <- bs(training$age,df=3)

lm1 <- lm(wage ~ bsBasis,data=training)
plot(training$age,training$wage,pch=19,cex=0.5)
points(training$age,predict(lm1,newdata=training),col="red",pch=19,cex=0.5)

```
Level 1 - Raw data to covariates
Level 2 - Covariates to covariates
 










