---
title: "Barbell lifts"
output: html_notebook
---

# Challenge

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. 

They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


```{r}
# clear
rm(list = ls())

# library
library(caret)
library(RANN)
library(dplyr)
# data
fullTraining = read.csv("/Users/admin/Git/marto12.github.io/coursera-sandbox/pml-training.csv",header = TRUE,na.strings = c("NA","NaN","","#DIV/0!"))
#finalTesting = read.csv("/Users/admin/Git/marto12.github.io/coursera-sandbox/pml-testing.csv")

```

# Split into test and train
```{r}
inTrain = createDataPartition(fullTraining$classe, p = 0.6)[[1]]
training = fullTraining[ inTrain,]
validation = fullTraining[-inTrain,]

# Test for missing values
# pc <- length(na.omit(training$raw_timestamp_part_1)) /length(na.omit(training$classe))

```

# Remove empty columns

```{r}
emptycols <- sapply(training, function (k) all(is.na(k)))
training <- training[!emptycols]

#training[, colSums(abs(training)) ! == 0]

```

# Imputate missing values with KNN 

```{r}
preProcValues <- preProcess(training, method = c("knnImpute")) # have to apply this to testing set too. 
training2 <- predict(preProcValues, training)
```

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

# Train Boosted Model 
```{r}
#trainBoosted <- train(classe~.,data=training2,method="gbm")
train <- train(classe~.,data=training2,method="glm")

```


```{r}
trainBoosted
trainSVM


# SVM 
set.seed(325)
library(e1071)
trainSVM <- svm(CompressiveStrength ~ ., data = training)
predictSVM  <- predict(trainSVM, testing[,-9])
accuracy(predictSVM,testing$CompressiveStrength)
```

